---
title: "Manual Annotation"
author:
- Joseph Bergenstråhle, Royal Institute of Technology (KTH)
- Ludvig Larsson, Royal Institute of Technology (KTH)
date: ''
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Info

As default, all capture-spots are annotated as "Default". By using the annotation tool, the user puts labels on the selected capture-spots. After closing the tool, the capture-spot annotatons are update accordingly. 

## Usage

Required: images and spotfile loaded prior to usage.
The tool opens the shiny application in the default browser. Follow the instructions in the sidepanel. 
After closing the tool, the seurat is updates with the included annotations. 

```{r, eval=FALSE, echo=TRUE}
se <- ST.annotation(object=se)
```

## Demonstration of how the annotations can be used for DEA

Lets use our tool to conduct a simple DEA. First, open the tool and make two selections. Lets call these "one" and "two". Now, we can perform DEA between the capture-spots annotated as "one" vs the capture-spots annotated as "two":

First we add the labels to Seurat identity class 

```{r, eval=FALSE, echo=TRUE}

#OBS need to double check if this overwrite old Idents
Idents(object = se) <- se$meta.data$labels

```

Then we can use `FindMarkers` to perform DEA for the identity classes

```{r, eval=FALSE, echo=TRUE}

de.markers <- FindMarkers(se, ident.1 = "one", ident.2 = "two")
head(de.markers)
```


We can exlude `ident.2` to instead perform differential expression between one group vs all other capture-spots.

```{r}
de.markers <- FindMarkers(se, ident.1 = "one")
head(de.markers)
```



