---
title: "Manual Annotation"
author:
- Joseph Bergenstråhle, Royal Institute of Technology (KTH)
- Ludvig Larsson, Royal Institute of Technology (KTH)
date: ''
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, eval=TRUE, echo=FALSE}

load("~/Documents/temp_for_anno_vignette.RData")

```

## Info

As default, all capture-spots are annotated as "Default". By using the annotation tool, the user puts labels on the selected capture-spots. After closing the tool, the capture-spot annotations are updated accordingly. 

## Usage

Required: images and spotfile loaded prior to usage.
The tool opens the shiny application in the default browser. Follow the instructions in the side panel. 
After closing the tool, the seurat object is updated with the annotation (stored in object@meta.data$labels) 

By parameter `res` the user can choose the resolution of the image (higher = more computationally heavy).
Default is set to 1500px, lower this if loading takes to long or running into memory issued. 

```{r, eval=FALSE, echo=TRUE}
se <- ST.annotation(object=se)
```

## Demonstration of how the annotations can be used for DEA

Lets use our tool to conduct a simple DEA. First, open the tool and make two selections. Lets call these "one" and "two". Now, we can perform DEA between the capture-spots annotated as "one" vs the capture-spots annotated as "two":

First we add the labels to Seurat identity class 

```{r, eval=TRUE, echo=TRUE}

#OBS overwrite old Idents?
Seurat::Idents(object = se) <- se@meta.data$labels

```

Then we can use `FindMarkers` to perform DEA for the identity classes

```{r, eval=TRUE, echo=TRUE}

de.markers <- Seurat::FindMarkers(se, ident.1 = "one", ident.2 = "two")
head(de.markers)
```


We can exude `ident.2` to instead perform differential expression between one group vs all other capture-spots.

```{r}
de.markers <- Seurat::FindMarkers(se, ident.1 = "one")
head(de.markers)
```


See https://satijalab.org/seurat/v3.0/de_vignette.html for various pre-filter options that can be used to increase the speed of the DE test. 


